Do we need detailed demographic data to forecast plant population responses to climate change?

### 2015-01-22 manuscript/manuscript.Rmd
updated ms with new figs and methods text

We fit all models using a hierarchical Bayesian approach, which we describe in more detail below. However, for each vital rate statistical model we also define the likelihood model we use. For the likelihood models, Y is always the relevant vector of observations (e.g., whether a  genet survived [1] or not [0] from year $t$ to $t+1$).

@@

#### Vital rate models: quadrat level
The statistical approach used to model vital rates using aggregated data depends on the type of data collected. In our case, and as is often the case with census data, we have percent cover data (which can easily be transformed to proportion data, of course). We first considered fitting three vital rate models analagous to those we fit at the individual level: one for probability of extirpation within a quadrat (analagous to survival), one for cover change within a quadrat (analagous to growth), and one for probability of colonization within a quadrat (analagous to recruitment). However, within-quadrat extirpation and colonization events were rare in our time series (N=X and N=X, respectively). Given the broad spatial distribution of the quadrats we are studying, it is safe to assume that these events are in fact rare enough to be ignored for our purposes. So we constrained our statistical modeling of vital rates at the population level to change in percent cover within quadrats. For most of the remaining discussion of statistical modeling we refer to proportion data, which is simply percent data divided by 100.

An obvious choice for fitting a linear model to proportion data is beta regression because the support of the beta distribution is [0,1], not including true zeros. However, when we used fitted model parameters from a beta regression in quadrat-based population model the simulated population tended toward 100% cover for all species. We therefore chose a more-constrained modeling approach based on a truncated log-normal likelihood. The model is

\begin{align}
x_{jq,t+1} &= \gamma^{G}_{j,t} + \phi^{G}_{jQ} + \beta^{G}_{j,t}x_{jq,t} + \theta^{S}_{jk}C_{k,t} \\
Y^{G}_{jq,t+1} &\sim \text{LogNormal}(x_{jq,t+1}, \tau{j}) T(0,1)
\end{align}

At the quadrat level we defined three vital rates:

1. Probability of extirpation ($S$): the probability that, for a given species, a particular quadrat will go from non-zero cover at time $t$ to zero cover at time $t+1$.
2. Cover change ($G$): the change in percent cover from time $t$ to $t+1$ for a given species within a particular quadrat.
3. Probability of colonization ($R$): the probability that, for a give species, a particular quadrat will go from zero cover at time $t$ to non-zero cover at time $t+1$.

We retain the abbreviations $S$, $G$, and $R$ from the analogous processes at the individual level. The vital rate models at the quadrat level are all based on quadrat proportional cover. Also, at the quadrat level we do not need to explicitly include a density dependent term. Since we are modeling proportional cover, we essentially get density-dependence for "free" when proportional cover is included as a covariate. That is, density-dependence emerges because quadrats with low cover generally increase in cover whereas quadrats with high cover generally decrease in cover.

We modeled the probability of extirpation of species $j$ in quadrat $q$ from time $t$ to $t+1$ as:

\begin{align}
\text{logit}(S_{jq,t}) &= \gamma^{S}_{j} + \phi^{S}_{jQ} + \beta^{S}_{j}x_{jq,t} + \theta^{S}_{jk}C_{k,t} \\
Y^{S}_{jq,t} &\sim \text{Bernoulli}(S_{jq,t})
\end{align}

where all parameters are as in Eq. 1 except that $x$ is now quadrat proportional cover. Note, however, that we do not include year random effects on the intercept $\gamma$ or the slope term $\beta$ for quadrat proportional cover. The quadrat data is inherently more sparse than the individual data from which it is aggregated, and this is especially evident when modeling rare events like extirpation and colonization. Thus, when we tried to fit random year effects, those terms did not converge.

We modeled quadrat cover change ($G$) from time $t$ to $t+1$ as:

\begin{equation}
\text{logit}(x_{jq,t+1}) = \gamma^{G}_{j,t} + \phi^{G}_{jQ} + \beta^{G}_{j,t}x_{jq,t} + \theta^{S}_{jk}C_{k,t}
\end{equation}

where, in this case, we do include random year effects on the intercept $\gamma$ and the slope term $\beta$. For cover change we had enough data for those terms to converge. Note that our model for quadrat cover change uses a logit transformation to link the expected cover at $t+1$ ($x_{jq,t}$) to the linear predictors. We do so because during model fitting we use a beta likelihood since the data, proportional cover, is beta distributed. The beta likelihood requires shape ($\rho$) and rate ($\eta$) parameters that can be calculated using moment-matching:

\begin{align}
\rho_{jq,t+1} &= x_{jq,t+1}\tau_{j} \\
\eta_{jq,t+1} &= (1-x_{jq,t+1})\tau_{j}
\end{align}

with likelihood:

\begin{equation}
Y^{G}_{jq,t+1} \sim \text{Beta}(\rho_{jq,t+1}, \eta_{jq,t+1}).
\end{equation}

Finally, we modeled probability of colonization in quadrat $q$ by species $j$ from time $t$ to $t+1$ as:

\begin{align}
\text{logit}(R_{jq,t}) &= \gamma^{R}_{j} + \phi^{R}_{jQ} + \theta^{R}_{jk}C_{k,t} \\
Y^{R}_{jq,t} &\sim \text{Bernoulli}(R_{jq,t}).
\end{align}

### Model fitting
Our Bayesian approach to fitting the vital rate models required choosing appropriate priors for unknown parameters and deciding which, if any, of those prior should be hierarchical. We decided to fit models where all terms except climate covariates were fit by species, while the climate covariates were fit hierarchically where species-specific coefficients were drawn from a shared `global' coefficient distribution. We did so for two reasons: (1) the four focal species are all perennial grasses that we expect to respond similarly to climate covariates, and (2) convergence of climate effects at the quadrat level was much easier to achieve when we modeled these terms hierarchically, allowing them to "share" statistical strength via partial pooling [@Gelman2007]. So, climate effects were modeled as:

@@

All of our analyses (model fitting and simulating) were conducted in R [@Team2013]. We used the MCMC sampler in JAGS [@Plummer2003] to estimate the posterior distributions of model parameters and the package 'r2jags' [@Su2012] to connect R to JAGS. We obtained posterior distributions for all model parameters from three parallel MCMC chains run for 50,000 iterations, after discarding an initial 50,000 iterations. We assessed convergence visually and using the @Gelman1992 diagnostic in the R package `coda' [@Plummer2006]. Scale reduction factors for all parameters were less than 1.02, indicating convergence. For the purposes of introducing stochasticity in our population models, we saved the final 1,000 iterations from each chain for all parameters to be used as randomly drawn values during population simulation.

We assessed the statistical importance of the climate covariates included the final vital rate regressions by comparing the residual deviance of models with climate covariates and temporal random effects, climate covariates only, and temporal random effects only. When a model includes climate covariates, this comparison shows the relative contribution of the climate covariates in explaining the total interannual variability [@Adler2012].

@@

We used random draws from the final 1,000 iterations from each of 3 MCMC chains to introduce stochasticity into our population models. At each time step, we first randomly selected climate covariates from one of the 14 observed years and also randomly drew one set of random year effects for each vital rate regression. Climate years and random year effects were drawn independently because they were uncorrelated. Then, we drew the full parameter set (specific random year effects, climate effects, fixed effects) from a randomly selected MCMC iteration. We selected random year effects and the climate year independently.

Results
-------
### Climate effects on vital rates
At the individual level we fit vital rate models that included climate covariates and random year effects (full model), just climate covariates (climate model), and no climate or random year effects (constant model). We used the metric from @Adler2012 to quantify the contribution of climate covariates to explaining interannual variability in vital rates based on comparisons of model deviance: (Climate model - Constant Model)/(Full model - Constant model). At the quadrat level, we could only fit random year effects for the growth regression. This is, in fact, one of the limitation of census-based data: local extirpation and colonization events are likely to be rare, making it difficult to fit complex statistical models. So, in this section we focus on growth only.

Climate covariates improved the growth regressions at both levels, with a slightly greater contribution at the individual level (Figure 1). Parameter estimates were generally consistent in terms of sign and magnitude between the individual level and the quadrat level for each species and climate covariate (Figure 2). There is no clear pattern suggesting that parameters estimated using individual level data are more certain, as evidenced by credible interval widths, than estimates based on quadrat data.

### Climate effects on population dynamics

@@

# ggplot(diffD, aes(color=simtype))+
#   geom_hline(aes(yintercept=0), linetype=2, color="grey50")+
#   geom_point(aes(x=as.character(year), y=avgDiff), size=3, position=position_dodge(width=0.5))+
#   geom_errorbar(aes(x=as.character(year), ymax=upDiff95, ymin=dnDiff95), width=0.00001, position=position_dodge(width=0.5))+
#   geom_errorbar(aes(x=as.character(year), ymax=upDiff75, ymin=dnDiff75), size=1, width=0.00001, position=position_dodge(width=0.5))+
#   facet_grid(species~., scales = "free_y")+
#   xlab("Year")+
#   ylab("Residuals (% cover difference)")+
#   theme_bw()+
#   scale_color_manual(values=myCols)+
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))
# 
# ggplot(resDall)+
#   geom_hline(aes(yintercept=0), linetype=2)+
# #   geom_line(data=diffD, aes(x=year, y=avgDiff, group=species), color="purple", alpha=0.5)+
# #   stat_smooth(data=diffD, aes(x=year, y=avgDiff, group=species), color="lightblue", se = FALSE, size=1)+
#   geom_boxplot(aes(x=year, y=resids, fill=simtype), outlier.shape = 1, width=0.3, outlier.size = 0.5)+
#   facet_grid(species~., scales = "free_y")+
#   xlab("Year")+
#   ylab("Residuals (% cover difference)")+
#   theme_bw()+
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))
# myCols <- c("#237DA4", "#7ECAD0")
p1 <- ggplot(subset(resDall, species=="BOGR"))+
  geom_hline(aes(yintercept=0), linetype=2)+
  geom_boxplot(aes(x=as.character(year), y=resids, fill=simtype), outlier.shape = 1, outlier.size = 0, color="grey25")+
  xlab("Year")+
  ylab("Residuals (% cover difference)")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_manual(values=myCols)+
  coord_cartesian(ylim = c(-50, 50))+
  ggtitle(expression(italic("B. gracilis")))+
  guides(fill=FALSE)
p2 <- ggplot(subset(resDall, species=="HECO"))+
  geom_hline(aes(yintercept=0), linetype=2)+
  geom_boxplot(aes(x=as.character(year), y=resids, fill=simtype), outlier.shape = 1, outlier.size = 0, color="grey25")+
  xlab("Year")+
  ylab("Residuals (% cover difference)")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_manual(values=myCols, labels=c("Year effects", "No year effects"))+
  ggtitle(expression(italic("H. comita")))+
  coord_cartesian(ylim = c(-10, 5))+
  guides(fill=guide_legend(title=""))+
  theme(legend.position=c(0.5,0.2))
p3 <- ggplot(subset(resDall, species=="PASM"))+
  geom_hline(aes(yintercept=0), linetype=2)+
  geom_boxplot(aes(x=as.character(year), y=resids, fill=simtype), outlier.shape = 1, outlier.size = 0, color="grey25")+
  xlab("Year")+
  ylab("Residuals (% cover difference)")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_manual(values=myCols)+
  ggtitle(expression(italic("P. smithii")))+
  guides(fill=FALSE)+
  coord_cartesian(ylim = c(-1, 1))
p4 <- ggplot(subset(resDall, species=="POSE"))+
  geom_hline(aes(yintercept=0), linetype=2)+
  geom_boxplot(aes(x=as.character(year), y=resids, fill=simtype), outlier.shape = 1, outlier.size = 0, color="grey25")+
  xlab("Year")+
  ylab("Residuals (% cover difference)")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_manual(values=myCols)+
  ggtitle(expression(italic("Poa secunda")))+
  guides(fill=FALSE)+
  coord_cartesian(ylim = c(-10, 15))
pFinal <- grid.arrange(p1,p2,p3,p4, ncol=2, nrow=2)
pFinal

@@

```{r figure_2, dependson="plot-options", fig.cap="Simulated changes species' median cover caused by a 1% increase in observed precipitation or temperature. We used the median of simulated results because the average was highly influence by some extreme outliers.", fig.width=4.5, fig.height=4, cache=FALSE, results='hide'}
climBase <- readRDS("../montana/quadBM/simulations/baselineSimulationQBM.rds")
climPpt <- readRDS("../montana/quadBM/simulations/pptChangeSimulationQBM.rds")
climTemp <- readRDS("../montana/quadBM/simulations/tempChangeSimulationQBM.rds")
climBoth <- readRDS("../montana/quadBM/simulations/temp-pptChangeSimulationQBM.rds")
climBase$sim <- rep("AObserved", nrow(climBase)) 
climPpt$sim <- rep("BPrecipitation", nrow(climBase)) 
climTemp$sim <- rep("CTemperature", nrow(climBase)) 
climBoth$sim <- rep("DBoth", nrow(climBase)) 
climSimsQBM <- rbind(climBase, climPpt, climTemp, climBoth)

>>>>

We fit all models using a hierarchical Bayesian approach, which we describe in more detail below. However, for each vital rate statistical model we also define the likelihood model we use. For the likelihood models, \textbf{Y} is always the relevant vector of observations (e.g., whether a  genet survived [1] or not [0] from year $t$ to $t+1$).

@@

#### Vital rate models: quadrat level
The statistical approach used to model vital rates using aggregated data depends on the type of data collected. In our case, and as is often the case with census data, we have percent cover data (which can easily be transformed to proportion data, of course). We first considered fitting three vital rate models analagous to those we fit at the individual level: one for probability of extirpation within a quadrat (analagous to survival), one for cover change within a quadrat (analagous to growth), and one for probability of colonization within a quadrat (analagous to recruitment). However, within-quadrat extirpation and colonization events were rare in our time series ($N=9$ and $N=10$, respectively). Given the broad spatial distribution of the quadrats we are studying, it is safe to assume that these events are in fact rare enough to be ignored for our purposes. So we constrained our statistical modeling of vital rates at the population level to change in percent cover within quadrats. For most of the remaining discussion of statistical modeling we refer to proportion data, which is simply percent data divided by 100.

An obvious choice for fitting a linear model to proportion data is beta regression because the support of the beta distribution is [0,1], not including true zeros or ones. However, when we used fitted model parameters from a beta regression in quadrat-based population model the simulated population tended toward 100% cover for all species. We therefore chose a more constrained modeling approach based on a truncated log-normal likelihood. The model for quadrat cover change ($G$) from time $t$ to $t+1$

\begin{align}
x_{jq,t+1} &= \gamma^{G}_{j,t} + \phi^{G}_{jQ} + \beta^{G}_{j,t}x_{jq,t} + \theta^{S}_{jk}C_{k,t} \\
Y^{G}_{jq,t+1} &\sim \text{LogNormal}(x_{jq,t+1}, \tau{j}) T[0,1]
\end{align}

where $x_{jq,t}$ is the log of species' $j$ proportional cover in quadrat $q$ at time $t$ and all other parameters are as in the individual-level growth model (Eq. #). The log normal likelihood includes a truncation ($T[0,1]) to ensure that predicted values do not exceed 100% cover.

### Model fitting
Our Bayesian approach to fitting the vital rate models required choosing appropriate priors for unknown parameters and deciding which, if any, of those priors should be hierarchical. We decided to fit models where all terms except climate covariates were fit by species, while the climate covariates were fit hierarchically with species-specific coefficients drawn from a shared `global' coefficient distribution. We did so for two reasons: (1) the four focal species are all perennial grasses that we expect to respond similarly to climate covariates, and (2) convergence of climate effects at the quadrat level was much easier to achieve when we modeled these terms hierarchically, allowing them to "share" statistical strength via partial pooling [@Gelman2007]. So, climate effects were modeled as:

@@

All of our analyses (model fitting and simulating) were conducted in R [@Team2013]. We used the MCMC sampler in JAGS [@Plummer2003] to estimate the posterior distributions of model parameters and the package 'r2jags' [@Su2012] to connect R to JAGS. We obtained posterior distributions for all model parameters from three parallel MCMC chains run for 50,000 iterations, after discarding an initial 50,000 iterations. We assessed convergence visually and using the @Gelman1992 diagnostic in the R package `coda' [@Plummer2006]. Scale reduction factors for all parameters were less than 1.01, indicating convergence. For the purposes of introducing stochasticity in our population models, we saved the final 1,000 iterations from each of the three MCMC chains for all parameters to be used as randomly drawn values during population simulation.

@@

We used random draws from the final 1,000 iterations from each of three MCMC chains to introduce stochasticity into our population models. At each time step, we first randomly selected climate covariates from one of the 14 observed years and also randomly drew one set of random year effects for each vital rate regression. Climate years and random year effects were drawn independently because they were uncorrelated. Then, we drew the full parameter set (specific random year effects, climate effects, fixed effects) from a randomly selected MCMC iteration. We selected random year effects and the climate year independently.

Results
-------

@@

# ggplot(diffD, aes(color=simtype))+
#   geom_hline(aes(yintercept=0), linetype=2, color="grey50")+
#   geom_point(aes(x=as.character(year), y=avgDiff), size=3, position=position_dodge(width=0.5))+
#   geom_errorbar(aes(x=as.character(year), ymax=upDiff95, ymin=dnDiff95), width=0.00001, position=position_dodge(width=0.5))+
#   geom_errorbar(aes(x=as.character(year), ymax=upDiff75, ymin=dnDiff75), size=1, width=0.00001, position=position_dodge(width=0.5))+
#   facet_grid(species~., scales = "free_y")+
#   xlab("Year")+
#   ylab("Residuals (% cover difference)")+
#   theme_bw()+
#   scale_color_manual(values=myCols)+
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))
# 
# ggplot(resDall)+
#   geom_hline(aes(yintercept=0), linetype=2)+
# #   geom_line(data=diffD, aes(x=year, y=avgDiff, group=species), color="purple", alpha=0.5)+
# #   stat_smooth(data=diffD, aes(x=year, y=avgDiff, group=species), color="lightblue", se = FALSE, size=1)+
#   geom_boxplot(aes(x=year, y=resids, fill=simtype), outlier.shape = 1, width=0.3, outlier.size = 0.5)+
#   facet_grid(species~., scales = "free_y")+
#   xlab("Year")+
#   ylab("Residuals (% cover difference)")+
#   theme_bw()+
#   theme(axis.text.x = element_text(angle = 45, hjust = 1))
# myCols <- c("#237DA4", "#7ECAD0")
p1 <- ggplot(subset(resDall, species=="BOGR"))+
  geom_hline(aes(yintercept=0), linetype=2)+
  geom_boxplot(aes(x=as.character(year), y=resids, fill=simtype), outlier.shape = 1, outlier.size = 0, color="grey25")+
  xlab("Year")+
  ylab("Residuals (% cover difference)")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_manual(values=myCols)+
  coord_cartesian(ylim = c(-50, 50))+
  ggtitle(expression(italic("B. gracilis")))+
  guides(fill=FALSE)
p2 <- ggplot(subset(resDall, species=="HECO"))+
  geom_hline(aes(yintercept=0), linetype=2)+
  geom_boxplot(aes(x=as.character(year), y=resids, fill=simtype), outlier.shape = 1, outlier.size = 0, color="grey25")+
  xlab("Year")+
  ylab("Residuals (% cover difference)")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_manual(values=myCols, labels=c("Year effects", "No year effects"))+
  ggtitle(expression(italic("H. comita")))+
  coord_cartesian(ylim = c(-10, 5))+
  guides(fill=guide_legend(title=""))+
  theme(legend.position=c(0.5,0.2))
p3 <- ggplot(subset(resDall, species=="PASM"))+
  geom_hline(aes(yintercept=0), linetype=2)+
  geom_boxplot(aes(x=as.character(year), y=resids, fill=simtype), outlier.shape = 1, outlier.size = 0, color="grey25")+
  xlab("Year")+
  ylab("Residuals (% cover difference)")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_manual(values=myCols)+
  ggtitle(expression(italic("P. smithii")))+
  guides(fill=FALSE)+
  coord_cartesian(ylim = c(-1.5, 1.5))
p4 <- ggplot(subset(resDall, species=="POSE"))+
  geom_hline(aes(yintercept=0), linetype=2)+
  geom_boxplot(aes(x=as.character(year), y=resids, fill=simtype), outlier.shape = 1, outlier.size = 0, color="grey25")+
  xlab("Year")+
  ylab("Residuals (% cover difference)")+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_manual(values=myCols)+
  ggtitle(expression(italic("Poa secunda")))+
  guides(fill=FALSE)+
  coord_cartesian(ylim = c(-10, 15))
pFinal <- grid.arrange(p1,p2,p3,p4, ncol=2, nrow=2)
pFinal

@@

```{r figure_2, dependson="plot-options", fig.cap="Simulated changes species' median cover caused by a 1% increase in observed precipitation or temperature. We used the median of simulated results because the average was highly influence by some extreme outliers.", fig.width=6.5, fig.height=4, cache=FALSE, results='hide'}
climBase <- readRDS("../montana/quadBM/simulations/baselineSimulationQBM.rds")
climPpt <- readRDS("../montana/quadBM/simulations/pptChangeSimulationQBM.rds")
climTemp <- readRDS("../montana/quadBM/simulations/tempChangeSimulationQBM.rds")
climBoth <- readRDS("../montana/quadBM/simulations/temp-pptChangeSimulationQBM.rds")
climBase$sim <- rep("AObserved", nrow(climBase)) 
climPpt$sim <- rep("BPrecipitation", nrow(climBase)) 
climTemp$sim <- rep("CTemperature", nrow(climBase)) 
climBoth$sim <- rep("DBoth", nrow(climBase)) 
climSimsQBM <- rbind(climBase, climPpt, climTemp, climBoth)

### 2015-01-22 manuscript/manuscript.tex
updated ms with new figs and methods text

We fit all models using a hierarchical Bayesian approach, which we
describe in more detail below. However, for each vital rate statistical
model we also define the likelihood model we use. For the likelihood
models, Y is always the relevant vector of observations (e.g., whether a
genet survived {[}1{]} or not {[}0{]} from year $t$ to $t+1$).

@@

The statistical approach used to model vital rates using aggregated data
depends on the type of data collected. In our case, and as is often the
case with census data, we have percent cover data (which can easily be
transformed to proportion data, of course). We first considered fitting
three vital rate models analagous to those we fit at the individual
level: one for probability of extirpation within a quadrat (analagous to
survival), one for cover change within a quadrat (analagous to growth),
and one for probability of colonization within a quadrat (analagous to
recruitment). However, within-quadrat extirpation and colonization
events were rare in our time series (N=X and N=X, respectively). Given
the broad spatial distribution of the quadrats we are studying, it is
safe to assume that these events are in fact rare enough to be ignored
for our purposes. So we constrained our statistical modeling of vital
rates at the population level to change in percent cover within
quadrats. For most of the remaining discussion of statistical modeling
we refer to proportion data, which is simply percent data divided by
100.

An obvious choice for fitting a linear model to proportion data is beta
regression because the support of the beta distribution is {[}0,1{]},
not including true zeros. However, when we used fitted model parameters
from a beta regression in quadrat-based population model the simulated
population tended toward 100\% cover for all species. We therefore chose
a more-constrained modeling approach based on a truncated log-normal
likelihood. The model is

\begin{align}
x_{jq,t+1} &= \gamma^{G}_{j,t} + \phi^{G}_{jQ} + \beta^{G}_{j,t}x_{jq,t} + \theta^{S}_{jk}C_{k,t} \\
Y^{G}_{jq,t+1} &\sim \text{LogNormal}(x_{jq,t+1}, \tau{j}) T(0,1)
\end{align}

At the quadrat level we defined three vital rates:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Probability of extirpation ($S$): the probability that, for a given
  species, a particular quadrat will go from non-zero cover at time $t$
  to zero cover at time $t+1$.
\item
  Cover change ($G$): the change in percent cover from time $t$ to $t+1$
  for a given species within a particular quadrat.
\item
  Probability of colonization ($R$): the probability that, for a give
  species, a particular quadrat will go from zero cover at time $t$ to
  non-zero cover at time $t+1$.
\end{enumerate}

We retain the abbreviations $S$, $G$, and $R$ from the analogous
processes at the individual level. The vital rate models at the quadrat
level are all based on quadrat proportional cover. Also, at the quadrat
level we do not need to explicitly include a density dependent term.
Since we are modeling proportional cover, we essentially get
density-dependence for ``free'' when proportional cover is included as a
covariate. That is, density-dependence emerges because quadrats with low
cover generally increase in cover whereas quadrats with high cover
generally decrease in cover.

We modeled the probability of extirpation of species $j$ in quadrat $q$
from time $t$ to $t+1$ as:

\begin{align}
\text{logit}(S_{jq,t}) &= \gamma^{S}_{j} + \phi^{S}_{jQ} + \beta^{S}_{j}x_{jq,t} + \theta^{S}_{jk}C_{k,t} \\
Y^{S}_{jq,t} &\sim \text{Bernoulli}(S_{jq,t})
\end{align}

where all parameters are as in Eq. 1 except that $x$ is now quadrat
proportional cover. Note, however, that we do not include year random
effects on the intercept $\gamma$ or the slope term $\beta$ for quadrat
proportional cover. The quadrat data is inherently more sparse than the
individual data from which it is aggregated, and this is especially
evident when modeling rare events like extirpation and colonization.
Thus, when we tried to fit random year effects, those terms did not
converge.

We modeled quadrat cover change ($G$) from time $t$ to $t+1$ as:

\begin{equation}
\text{logit}(x_{jq,t+1}) = \gamma^{G}_{j,t} + \phi^{G}_{jQ} + \beta^{G}_{j,t}x_{jq,t} + \theta^{S}_{jk}C_{k,t}
\end{equation}

where, in this case, we do include random year effects on the intercept
$\gamma$ and the slope term $\beta$. For cover change we had enough data
for those terms to converge. Note that our model for quadrat cover
change uses a logit transformation to link the expected cover at $t+1$
($x_{jq,t}$) to the linear predictors. We do so because during model
fitting we use a beta likelihood since the data, proportional cover, is
beta distributed. The beta likelihood requires shape ($\rho$) and rate
($\eta$) parameters that can be calculated using moment-matching:

\begin{align}
\rho_{jq,t+1} &= x_{jq,t+1}\tau_{j} \\
\eta_{jq,t+1} &= (1-x_{jq,t+1})\tau_{j}
\end{align}

with likelihood:

\begin{equation}
Y^{G}_{jq,t+1} \sim \text{Beta}(\rho_{jq,t+1}, \eta_{jq,t+1}).
\end{equation}

Finally, we modeled probability of colonization in quadrat $q$ by
species $j$ from time $t$ to $t+1$ as:

\begin{align}
\text{logit}(R_{jq,t}) &= \gamma^{R}_{j} + \phi^{R}_{jQ} + \theta^{R}_{jk}C_{k,t} \\
Y^{R}_{jq,t} &\sim \text{Bernoulli}(R_{jq,t}).
\end{align}

@@

Our Bayesian approach to fitting the vital rate models required choosing
appropriate priors for unknown parameters and deciding which, if any, of
those prior should be hierarchical. We decided to fit models where all
terms except climate covariates were fit by species, while the climate
covariates were fit hierarchically where species-specific coefficients
were drawn from a shared `global' coefficient distribution. We did so
for two reasons: (1) the four focal species are all perennial grasses
that we expect to respond similarly to climate covariates, and (2)
convergence of climate effects at the quadrat level was much easier to
achieve when we modeled these terms hierarchically, allowing them to
``share'' statistical strength via partial pooling (Gelman and Hill
2007). So, climate effects were modeled as:

@@

All of our analyses (model fitting and simulating) were conducted in R
(Team 2013). We used the MCMC sampler in JAGS (Plummer 2003) to estimate
the posterior distributions of model parameters and the package `r2jags'
(Su and Yajima 2012) to connect R to JAGS. We obtained posterior
distributions for all model parameters from three parallel MCMC chains
run for 50,000 iterations, after discarding an initial 50,000
iterations. We assessed convergence visually and using the Gelman and
Rubin (1992) diagnostic in the R package `coda' (Plummer et al. 2006).
Scale reduction factors for all parameters were less than 1.02,
indicating convergence. For the purposes of introducing stochasticity in
our population models, we saved the final 1,000 iterations from each
chain for all parameters to be used as randomly drawn values during
population simulation.

We assessed the statistical importance of the climate covariates
included the final vital rate regressions by comparing the residual
deviance of models with climate covariates and temporal random effects,
climate covariates only, and temporal random effects only. When a model
includes climate covariates, this comparison shows the relative
contribution of the climate covariates in explaining the total
interannual variability (Adler et al. 2012).

@@

We used random draws from the final 1,000 iterations from each of 3 MCMC
chains to introduce stochasticity into our population models. At each
time step, we first randomly selected climate covariates from one of the
14 observed years and also randomly drew one set of random year effects
for each vital rate regression. Climate years and random year effects
were drawn independently because they were uncorrelated. Then, we drew
the full parameter set (specific random year effects, climate effects,
fixed effects) from a randomly selected MCMC iteration. We selected
random year effects and the climate year independently.

@@

\subsubsection{Climate effects on vital
rates}\label{climate-effects-on-vital-rates}

At the individual level we fit vital rate models that included climate
covariates and random year effects (full model), just climate covariates
(climate model), and no climate or random year effects (constant model).
We used the metric from Adler et al. (2012) to quantify the contribution
of climate covariates to explaining interannual variability in vital
rates based on comparisons of model deviance: (Climate model - Constant
Model)/(Full model - Constant model). At the quadrat level, we could
only fit random year effects for the growth regression. This is, in
fact, one of the limitation of census-based data: local extirpation and
colonization events are likely to be rare, making it difficult to fit
complex statistical models. So, in this section we focus on growth only.

Climate covariates improved the growth regressions at both levels, with
a slightly greater contribution at the individual level (Figure 1).
Parameter estimates were generally consistent in terms of sign and
magnitude between the individual level and the quadrat level for each
species and climate covariate (Figure 2). There is no clear pattern
suggesting that parameters estimated using individual level data are
more certain, as evidenced by credible interval widths, than estimates
based on quadrat data.

\subsubsection{Climate effects on population
dynamics}\label{climate-effects-on-population-dynamics}

@@

@@

Adler, P. B., H. J. Dalgleish, and S. P. Ellner. 2012. Forecasting plant
community impacts of climate variability and change: when do competitive
interactions matter? Journal of Ecology 100:478--487.

>>>>

We fit all models using a hierarchical Bayesian approach, which we
describe in more detail below. However, for each vital rate statistical
model we also define the likelihood model we use. For the likelihood
models, \textbf{Y} is always the relevant vector of observations (e.g.,
whether a genet survived {[}1{]} or not {[}0{]} from year $t$ to $t+1$).

@@

The statistical approach used to model vital rates using aggregated data
depends on the type of data collected. In our case, and as is often the
case with census data, we have percent cover data (which can easily be
transformed to proportion data, of course). We first considered fitting
three vital rate models analagous to those we fit at the individual
level: one for probability of extirpation within a quadrat (analagous to
survival), one for cover change within a quadrat (analagous to growth),
and one for probability of colonization within a quadrat (analagous to
recruitment). However, within-quadrat extirpation and colonization
events were rare in our time series ($N=9$ and $N=10$, respectively).
Given the broad spatial distribution of the quadrats we are studying, it
is safe to assume that these events are in fact rare enough to be
ignored for our purposes. So we constrained our statistical modeling of
vital rates at the population level to change in percent cover within
quadrats. For most of the remaining discussion of statistical modeling
we refer to proportion data, which is simply percent data divided by
100.

An obvious choice for fitting a linear model to proportion data is beta
regression because the support of the beta distribution is {[}0,1{]},
not including true zeros or ones. However, when we used fitted model
parameters from a beta regression in quadrat-based population model the
simulated population tended toward 100\% cover for all species. We
therefore chose a more constrained modeling approach based on a
truncated log-normal likelihood. The model for quadrat cover change
($G$) from time $t$ to $t+1$

\begin{align}
x_{jq,t+1} &= \gamma^{G}_{j,t} + \phi^{G}_{jQ} + \beta^{G}_{j,t}x_{jq,t} + \theta^{S}_{jk}C_{k,t} \\
Y^{G}_{jq,t+1} &\sim \text{LogNormal}(x_{jq,t+1}, \tau{j}) T[0,1]
\end{align}

where $x_{jq,t}$ is the log of species' $j$ proportional cover in
quadrat $q$ at time $t$ and all other parameters are as in the
individual-level growth model (Eq. \#). The log normal likelihood
includes a truncation (\$T{[}0,1{]}) to ensure that predicted values do
not exceed 100\% cover.

@@

Our Bayesian approach to fitting the vital rate models required choosing
appropriate priors for unknown parameters and deciding which, if any, of
those priors should be hierarchical. We decided to fit models where all
terms except climate covariates were fit by species, while the climate
covariates were fit hierarchically with species-specific coefficients
drawn from a shared `global' coefficient distribution. We did so for two
reasons: (1) the four focal species are all perennial grasses that we
expect to respond similarly to climate covariates, and (2) convergence
of climate effects at the quadrat level was much easier to achieve when
we modeled these terms hierarchically, allowing them to ``share''
statistical strength via partial pooling (Gelman and Hill 2007). So,
climate effects were modeled as:

@@

All of our analyses (model fitting and simulating) were conducted in R
(Team 2013). We used the MCMC sampler in JAGS (Plummer 2003) to estimate
the posterior distributions of model parameters and the package `r2jags'
(Su and Yajima 2012) to connect R to JAGS. We obtained posterior
distributions for all model parameters from three parallel MCMC chains
run for 50,000 iterations, after discarding an initial 50,000
iterations. We assessed convergence visually and using the Gelman and
Rubin (1992) diagnostic in the R package `coda' (Plummer et al. 2006).
Scale reduction factors for all parameters were less than 1.01,
indicating convergence. For the purposes of introducing stochasticity in
our population models, we saved the final 1,000 iterations from each of
the three MCMC chains for all parameters to be used as randomly drawn
values during population simulation.

@@

We used random draws from the final 1,000 iterations from each of three
MCMC chains to introduce stochasticity into our population models. At
each time step, we first randomly selected climate covariates from one
of the 14 observed years and also randomly drew one set of random year
effects for each vital rate regression. Climate years and random year
effects were drawn independently because they were uncorrelated. Then,
we drew the full parameter set (specific random year effects, climate
effects, fixed effects) from a randomly selected MCMC iteration. We
selected random year effects and the climate year independently.

@@

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-figure_2.pdf}
\caption{Simulated changes species' median cover caused by a 1\%
increase in observed precipitation or temperature. We used the median of
simulated results because the average was highly influence by some
extreme outliers.}
\end{figure}

@@
